---
title: "City Mobility Insight"
author: "Omar Ali Fdal"
date: "1 November 2017"
output: 
  slidy_presentation:
    theme: "journal"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(leaflet)
library(shiny)
library(dplyr)
library(lubridate)

timeFormat <- "%H%M"
tileUrl <- "https://cartodb-basemaps-{s}.global.ssl.fastly.net/light_all/{z}/{x}/{y}.png"

tt <- read.csv("triptemplates_20171027.csv")
tt$activeColor <- "blue"
tt$activeColor[tt$active=="false"] <- "gray"
tt <- mutate(tt,
               depDatetimeLocal=parse_date_time(depDatetimeLocal,
                                                "ymdHMS"),
               depTime=format(depDatetimeLocal, timeFormat))

```

## City Mobility Insight
City Mobility Insight is an app that allows a City Transportation Operator to see how are the city commuters distributed, both geographically and temporarilly.

### Context
This app was developed during the Developing Data Products Coursera Course.

### The Running App
You can have a look at the app on (https://oalifdal.shinyapps.io/cityMobility/)[https://oalifdal.shinyapps.io/cityMobility/]

### Data Source
The data used is based on an urban carpooling app user base.

## Data Description

- active: Boolean value whether a commuter is active or not.
- visibility: Categorical variable whether a commuter is driver or passenger.
- depDatetimeLocal: departure time for the commuter's regular trip.
- departureLat: departure latitude
- departureLng: departure longitude

## Map Example: 
### Paris Commuters between 7am and 8am
```{r paris, echo=FALSE}
bounds <- c(2.312082, 48.821456, 2.392082, 48.911456)

minTime <- "07:00"
maxTime <- "08:00"
filteredTt <- tt[minTime <= tt$depTime & tt$depTime <= maxTime, ]

leaflet(filteredTt) %>%
  fitBounds(bounds[1], bounds[2],
            bounds[3], bounds[4]) %>%
  addTiles(urlTemplate = tileUrl) %>%
  addCircleMarkers(lat=~departureLat,
                 lng=~departureLng,
                 weight=1,
                 color=tt$activeColor) %>%
  addLegend(position="bottomleft",
            labels=c("Active", "Inactive"),
            colors=c("blue", "gray"))
```

## Code Example: Reactive updates

### Define reactive variables
```{r cars, echo = TRUE, eval=FALSE}
minDepTime <- reactive({format(input$minDepTime, timeFormat)})
maxDepTime <- reactive({format(input$maxDepTime, timeFormat)})

filteredTt <- reactive({
  tt[minDepTime() <= tt$depTime & tt$depTime <= maxDepTime(),] 
})
```

### Update outputs based on change in inputs
```
  observe({
    newdata <- filteredTt()
    leafletProxy("tripMap", data=newdata) %>% 
      clearMarkers() %>% clearShapes() %>%
      addCircleMarkers(lat=~departureLat,
                       lng=~departureLng,
                       weight=1,
                       color=tt$activeColor)
  })
```





